The ``Dark Side'' of Phylogenetic Comparative Methods

### 2014-08-21 Manuscript/dark.tex
Added section on phylogenetic folklore

\section{Why aren't more methods-users aware of the ``dark side'' of PCMs?}

@@

Program defaults are another thing which become folklore although tey are often arbitrary.

@@

%Some stuff isn’t published, or is published and isn’t justified - “folklore” and rules of thumb (±3 from Jones&Purvis97).

%Notes on how citation counts were extracted:
%We extracted yearly citation counts from the Web of Science on 22nd May 2013 for the following papers: We also obtained citation counts for \citet{harvey1991comparative} using Google Scholar, as Web of Science does not provide citation counts for books.

\subsubsection{2. Literature is technical, dense and numerous}

\subsubsection{3. Important information is hard to locate}

\subsubsection{4. Users jump straight to the implementation of the method}

@@

\subsection{Recommendations}

@@

%Users need to take responsibility too. Work harder to understand. Read widely. Don't take results at face value. Become less trusting! Simulate data to check method does what you think it should do. Don't be afraid to question standard practice - sometimes it's just folklore.

>>>>

\section{Why don't all methods-users know about the ``Dark Side'' of PCMs?}

@@

A lot of the information needed to properly apply PCMs is not found in the literature. 
We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students and over time these myths and rules-of-thumb become set in stone. 
One example of such folklore is in the defaults of programs that perform PCMs. 
These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.

% Do we want a specific example here? Theta estimation in OU models? 
% Could mention ±3 from Jones&Purvis97 - totally unjustified rule of thumb now = canon.

Other information about the limitations of a method may be absent from the literature due to the time lag between a new method being published and others having time to test and critique it. 
For example, \citealp{felsenstein1985phylogenies} published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published %add citations.
This time lag is shorter with more recent methods because more simulations testing the method are now required by journals.

\subsection{2. Literature is technical, dense and numerous}

@@

%Some stuff isn’t published, or is published and isn’t justified - “folklore” and rules of thumb (

\subsection{3. Important information is hard to locate}

\subsection{4. Users jump straight to the implementation of the method}

@@

\section{How can we solve some of these problems?}

\subsection{1. Simplify, summarize and share}
\subsection{2. Read, think and learn}
\subsection{3. Collaboration}
\subsection{4. Incentives}

@@

%Users need to take responsibility too. Work harder to understand. Read widely. Don't take results at face value. Become less trusting! Simulate data to check method does what you think it should do. Don't be afraid to question standard practice - sometimes it's just folklore. My pglm vs pgls example.
